---
title: "SOC 280 Lab 1 - Installing R"
output:
  html_document:
    df_print: paged
---

Welcome to the lab portion of Sociology 280: Data and Society. Over the course of this
lab, we'll learn to use the R programming language to read, process, transform,
analyze, and visualize data. Each lab will have an Rmarkdown notebook, like this 
one, which you'll follow along with. The cool thing about Rmarkdown notebooks is
that they easily combine plain text, like this, and 'code chunks' where we can write,
execute, and examine R code. The notebooks have a combination of worked examples,
questions, and programming challenges. 

In this first lab, our main goal is to just install and set up R and R studio, the 
tools we'll use to write and process the code that we write. After that, this 
notebook is mostly an example of the full "data science pipeline" from asking a research
question to answering it with a data visualization. This is a _demo_, so don't worry
if you don't follow what's going on today. But you will learn to do everything
that I do today.


Today, we're going to start with a simple question: What was the 'best' movie in
each decade of the last 100 years?

We'll use data from IMDB, and consider 'best' to mean 'highest rated on IMDB'.

We'll go through a few steps:

1. Asking a research question (done!)
2. Preparing our coding environment
3. Loading our data
4. Understanding the format of our data
5. Transforming our data to address our research question
6. Using data analysis to answer our question
7. Creating a visualization to address our research question
8. Finalizing our work as an .html file

# 1. Asking a research question (done!)

We did this already! Nice.   
"What was the 'best' (based on IMDB) movie in each decade of the last 100 years?

# 2. Preparing our coding environment
R can do a lot right out of the box, but usually we want to do more than just the
'base' R abilities. We can load extra functions as 'packages' or 'libraries'. 

For this project, and for most of this course, we'll use a set of libraries called
the `tidyverse`. In order to load them, we'll use our first code chunk. Run the 
chunk by pressing the green 'play' arrow.

```{r a code chunk}
library(tidyverse)
```


# 3. Loading our data
Ok, now that our environment is set up, we can load our data. We do that using 
a function. A function is a bit of code that stands in for lots of other code. It
lets us do something complicated with only one line of code. A function has two 
basic parts: the function name, and the function arguments. The arguments are written
in parentheses after the function name and they're extra instructions we give to the
function. Below, I use the `read_csv` function to read our data. I use the assingment
opporator `<-` to assign the result of reading the data to the name 'movies'.

Run the code chunk

```{r loading data}
movies <- read_csv('data/movies.csv')
movies
```

You'll see that I've also outputted the top few rows and columns of our data.
Additionally, if you look over in the "Environment" pane, you'll see that there's
a new object there called 'movies' with 2665 obs. of 11 variables. This is a 
small portion of the full IMDB movie dataset which has millions of movies. I've
only included the movies which got most ratings (not necessarily the highest ratings)

# 4. Understanding the format of our data
Ok, so now we have some cool data. We want to understand the data format to see 
how we can answer our question of what the best movie from each decade is.

We'll use another function, the `glimse` function, to look at all of the columns
in our data so we can understand what we have.


```{r}
glimpse(movies)
```
Of our 11 columns, I can see 4 that I need to answer my question.

First, I need the `titleType` variable so that I can limit my analysis only to 
movies. Next, I need the `primaryTitle` variable so I can identify the moves. I
also need to know the decade the movie is in, so I need the `startYear` variable.
Finally, I want the highest rated movie from each decade, so I'll need to attend
to the `averageRating` variable.

Once we know which variables we'll need, it's a good idea to examine their descriptive
statistics. R has a great function for this, `summary`. But I don't want to use
`summary` on the whole dataset, just on the four columns. So I use a different function,
`select` to limit the columns. I also use a 'pipe' operator `|>` to move my data
between functions.

```{r}
movies %>% select(titleType,primaryTitle,startYear,averageRating) %>%
  summary()
```
We can see that we have movies from 1921 until 2023, with a minimum rating of
1.9 and a max of 10. The years also suggest that a large majority of movies are
from after 2000.

I also want to know how many of the titles aren't movies, so I'll count the values
in the `titleType` variable:

```{r}
movies %>% count(titleType)
```

Great, it looks like almost all of the values are Movies, with a few hundred TV
shows.

# 5. Transforming our data to address our research question

As things stand now, we can't answer our question. We're missing a variable for the
decade. But we have a variable for the year. So we need a way to make a new 
variable that calculates the decade based on the year.

I do that below by dropping the last digit from the year and adding a zero. The
mutate function allows me to make a new variable. We'll learn more about this soon.

We also want to limit our analysis only to movies. I do that below using the
`filter` function.

I assign the result of my transformation to a new object `movies_clean` which 
I can use for my analysis in the next step.

```{r}

movies_clean <- movies %>% mutate(decade = str_c(str_sub(startYear, 1,3),"0")) %>%
  filter(titleType == 'movie')

```


# 6. Using data analysis to answer our question

Now that we have the data transformed and filtered, we can try answering our question.
We want the best movie by decade, so we'll need to group the movies by decade and then
pick the one with the highest rating in each group. We should end up with just
11 movies - one for each decade from the 1920s to the 2020s.

I'll use a combination of three functions `group_by`, `mutate`, and `filter`.

```{r}
our_answer <- movies_clean %>% group_by(decade) %>% mutate(rank = row_number(desc(averageRating))) %>%
  filter(rank == 1) %>% select(decade, primaryTitle, averageRating, genres, everything()) %>%
  arrange(startYear)
our_answer
```
Ok! Well, there's our answer as a table. Let's also make a simple visualization.

# 7. Creating a visualization to address our research question
To make the visualization, we'll use an additional tidyverse library, ggplot. I'm
not sure what the gg stands for, but I like to think it means "good good plot".

```{r}
our_answer %>% ggplot(aes(decade, averageRating, label = primaryTitle))+
  geom_col(alpha = .4)+
  geom_text(angle = 90, nudge_y = -4.3, size = 3)+
  ggtitle("Best Movie From Each Decade", "Based on IMDB ratings")+
  xlab('Decade')+
  ylab('Rating')+
  theme_classic()
```

Awesome. I've made a pretty simple, but informative graph. It shows the name of 
the 'best' movie from each decade as well as the relative rankings. It seems like
the 70s and the 90s represented a bit of a filmic golden age.

To be clear, this analysis is acatully slightly _simpler_ than what I expect
on your final project, where you'll present 2 visualizations. But! Something like
this could be a good start~

# 8. Finalizing our work as an .html file
One last thing. A cool feature of an Rmarkdown document is that we can 'knit' it
into an .html file. This makes it easy to, say do your work in a document like this
and then automatically turn it into a professional-looking blog post. There are 
a lot of options for knitting that we can use, and we'll learn about those later.
For now, we just need to go up to the button where it says 'Preview' and select the 
dropdown for knit to HTML.



