---
title: "SOC 280 Lab Meeting 2"
output: html_notebook
---

# Part 0: How to R

http://r4ds.had.co.nz/


Let's start with the layout: source, console, environment, files.

Let's try some stuff in the console:

1 / 200 * 30

(59 + 73 + 2) / 3

sin(pi / 2)

Let's assign some variables

x <- 100

y <- 1:100

snake_case <- "An easy way to store variable names in R. All lowercase letters with underscores for spaces. You can put numbers on the end if you want"

You might also see:

camelCase which uses an uppercase letter to denote new words

Functions

mean()

rnorm()

R Markdown

Just like last week, this is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r}
library(tidyverse)
```

What happened?
If you try to load a package and you can't, it probably means it's not installed. We can install a package using install.packages('somePackage'). Try typing the code to install the package 'tidyverse' in the console.

Once we have the packages we need loaded, we can import the data.

```{r}
airbnb <- read_csv('data/airbnb.csv')
```

Notice that you now have a new variable in your environment pane.

What is the role of each part of this code?
`airbnb <- read_csv('data/airbnb.csv')`

airbnb:
read_csv:
'data/airbnb.csv':

Now you'll make a new code chunk. You can either type out ``` using the key above 'tab', then {r} or use the handy shortcut 'command + option + i' on mac or 'control + alt + i' on linux or windows. In the new chunk, use glimpse() to look at the airbnb data.



## Part 1: Different data formats, data structures


1.1. Use the output from the chunk you made above to answer these questions. What different types of data do you see in the airbnb data set? How do you think these data were collected by Airbnb? How did they come to us? Is this big-data based on the definition we've used in class?

1.2. Run the following code chunk and then view the resulting variables by typing their names into the console.

```{r}
var1 <- mean(airbnb$number_of_reviews, na.rm = TRUE)
var2 <- unique(airbnb$neighbourhood)
var3 <- airbnb %>% select(latitude,longitude)
```

1.3. What are the differences between var1, var2, and var3 in terms of the kind of data and the format or structure it's stored in?

1.4. It's not a good idea to call things 'var1' or 'var2.' Based on what is stored by those variables, rename them using snake_case.

Bonus 1.1: What's the most common type of data in airbnb?

Bonus 1.2: Plot latitude and longitude, more bonus: use 'neighborhood' as the color most bonus: put it on a map


## Part 2: Introduction to the Variables, R environment, Data Structures Types

2.1 What variables have we seen already?

2.2 What functions have we seen already?

2.3 Use the function 'class()' to check the data type of the three variables from part 1.

### Varibles

In R we can store most data, from a single number to a huge spreadsheet, as a 'variable'.
These variables are basically a way of giving some peice of data a name so that
we can access it and easily use it later. You've already created some variables--
var1, var2, var3 are all variables, as are your re-named versions. Even `airbnb` which
is a dataframe with almost 9k observations is a variable. 

Here's an example of a variable:

```{r}
ians_full_name <- 'Ian Taylor Kennedy'
```

I wrote a variable name in snake_case 'ians_full_name' assigned it the value of my 
full name 'Ian Taylor Kennedy'

Technically, variables are one type of 'object' that R is worried about. The other main type 
of object that we'll deal with are functions, which we'll spend more time on in part 3.

2.4 Make a new chunk below. Use it to make a new variable that has your name in it.

### R environment

The R environment is all of the variables and functions available to you. You can 
see the variables loaded into your environment in the environment pane. You can
see the functions and other objects (including built-in data!) that are available by 
selecting packages from the dropdown menu labeled "Global Environment".

2.5 Look at the data that's available from the package 'datasets'. Write one of 
the available dataset names here:


### Data Types and Structures


You've probably noticed that we've stored a variety of stuff as variables. 
var1 is a single number
var2 is has 77 neighborhoods
var3 has 8949 pairs of numbers
ians_full_name has one name
airbnb has 8949 observations of 18 variables!

These variables reflec different data types and data structures .

*Data types* refer to the way data is stored, but we can think about it as what kind
of data is being stored. We can have numeric data (like var1), character data 
(var2, ians_full_name), but also special data types for dates, for large floating point numbers
for spatial geometry, and for other specialized uses.


*Data structures* refer to the way data is organized. In R the basic data structure
is a 'vector', which is an ordered set of data that's all the same type. You can
make a vector in R by using the c() function.

```{r}
c(2,1,2,55,3,25,6,5)
```


You can also make lists (using the list() function), which -aren't- all of the 
same type. Notice that I have three vectors in this list!

```{r}
list(var1, var2, ians_full_name)
```


DataFrames are key, we'll use them a lot. DataFrames are special lists of vectors where each of the vectors has the same length--these are the columns. Each column contains a specific kind of data. A row is the value in each column for a particular index level. 

We can get specific columns from a dataframe by using a $:
Or by asking for its name in []

```{r}
airbnb$room_type # returns a vector
airbnb['room_type'] # returns a dataframe with one column
```


Columns are vectors, so you can find a specific index by using []:

```{r}
airbnb$room_type[30]
```

You can use the function head(n) to get the top n values of a dataframe or list

```{r}
head(airbnb$room_type)
```


'room_type' has a few different values. We can use the function unique() to find the different values.

```{r}
unique(airbnb$room_type)
```


Try running each line of the code chunk below by putting your cursor on the line and pressing 'command' + 'return' on mac and 'ctl'+'return' on windows or linux.

```{r}
airbnb$room_type
airbnb$room_type[30]
head(airbnb$room_type)
unique(airbnb$room_type)
```


## 3 Functions
We can think about R functions as verbs, and variables as nouns. All R functions have the following form:

functionName(argument1 = value1, argument2 = value2)

You can see help information about any function by using the help() function, like

help('setwd') 

BUT there's a shortcut:

?setwd

OR if you're not sure of the name of the function:

??select

If none of that helps, remember the secret to coding: google and stackexchange

Let's try it out with a new function for dataframes, called 'count()'. Count let's
us see how many of each different value there is in the data. We can write it like this:

count(our_data, column_to_count)

Now try it using the airbnb data to count how many of each room types there are.
So make a new chunk, then use 'airbnb' as the name of our dataframe and 'room_type'
as our column name.



## 4 Practice Exercises

4.1 Make a new code chunk and use it to look at the summary of each column in airbnb using the function `summary()`, which takes a dataframe as its argument

4.2 Use the unique function to list the unique property names. How many are there?

4.3 find the average number of days that a property is available over the next year. You'll need to find the right variable, then the right function to use, then apply that function to that variable.

Bonus 4.1: What neighborhood in Chicago has the most airbnb listings?

Bonus 4.2: What's the average price of a unit?

