---
title: "SOC 280 Lab Workshop 4"
output: html_notebook
---

Check-In: 

# Part 0: Review

0.1 In this lab we'll be working with data about neighborhoods and health. The 
data come from two sources, the Places Nanda Survey, which surveys people about
their health, and the US Census's ACS (American Community Survey) from 2013-2017.
These data are at the census tract level, so each row represents a Census Tract.
The data are on blackboard in a file called 'health_census.csv'. Download the file 
and put it in your data folder.

0.2 Load the tidyverse and the health_census data. 

0.3 Glimpse the health_census data

# Part 1:

We'll start with an example:

Some questions are best answered visually. Let's use this one:

What is the relationship between neighborhood poverty and rates of Asthma.
*take a moment to make a hypothesis*

First, without a visualizaiton, I could look at a correlation:
```{r}
# some variables have missing values, so we have to drop those for this part

health_census_clean <- health_census %>% 
  select(asthma, cancer, ppov13_17, affluence13_17,pnhwhite13_17, totpop13_17) %>% 
  drop_na()
# do a simple correlation between my variables

cor(health_census_clean$ppov13_17, health_census_clean$asthma)
```
This result shows that poverty and asthma are positively correlated. In other words,
neighborhoods with more poverty also tend to have more asthma. Why would that be? 

How could we answer this question using visualizations? 

In this situation, though, we still don't have a good understanding of this relationship. It might help to make a graph.


```{r}
#OK a lot going on here
#* First we use the ggplot function and pass it two arguments
#* data, which tells it what data to use
#* and mapping, which tells it what variables to put where
#* Then we add a '+' to say, we're going tell you more about this plot on the next line
#* On the next line we just put 'geom_point' to tell ggplot to make a scatterplot

ggplot(data = health_census_clean,mapping =aes(x = ppov13_17, y = asthma)) + 
  geom_point()
```

1.1 What is the correlation between affluence and cancer?

1.2 Plot the relationship.

1.Bonus: Graph the relationship between price and number of characters.

# Part 2: Research Questions and Data Sources

What makes a good visualization?

Check out these resources to answer the following questions.

The Double Think: https://thedoublethink.com/tuftes-principles-for-visualizing-quantitative-information/

Darkhorse: https://www.darkhorseanalytics.com/blog/data-looks-better-naked

2.1 What is the data-ink ratio?

2.2 What is chart junk?

2.3 Look at the example chart on the darkhorse website. Use some specific examples to describe how they improve their visualization.

2.4 Which change from 1.3 do you think is most effective?

2.Bonus Modify my code from Part 1 to make a better visualization.



# Part 3: Adding on to visualizations

Let's consider our square footage graph from before:

```{r}
ggplot(data = health_census_clean,mapping =aes(x = ppov13_17, y = asthma)) + 
  geom_point()
```

We might be interested in how race interacts with tis.

```{r}
ggplot(data = health_census_clean,mapping =aes(x = ppov13_17, y = asthma, color = pnhwhite13_17)) + 
  geom_point()
```

What does this plot suggest about the relationship between asthma, poverty, and race?

We can also change the size of the dots. Now they'll be bigger when the total population
of the area is higher. I've changed the 'alpha' value in geom_point, which makes our
dots slightly see-through and lets us see more of them.

```{r}
ggplot(data = health_census_clean,mapping =aes(x = ppov13_17, y = asthma, size = totpop13_17)) + 
  geom_point(alpha = .1)
```



We can also add trendlines:
```{r}
ggplot(data = health_census_clean,mapping =aes(x = ppov13_17, y = asthma)) + 
  geom_point()+
  geom_smooth(method = 'lm')
```

There are some other kinds of plots we can make:

### Histograms:

Histograms show the variation within a particular variable.

```{r}
ggplot(health_census_clean, aes(asthma))+
  geom_histogram()
```


### Box and whisker plots:

Box and whisker plots are a good way to see distributions by some other variable.

In this case, we'll make three groups and see how asthma shows up in those groups.


```{r}

# here I make a new varible with mutate that has three groups for low, middle,
# and high poverty
health_census_clean <- health_census_clean %>% 
  mutate(pov_cat = cut_interval(ppov13_17, n = 3, labels = c('low poverty', 'middle poverty', 'high poverty')))

ggplot(health_census_clean, aes(pov_cat, asthma))+
  geom_boxplot()
```

### Bar graphs:

Let's make a bar graph of the number of observations we have in each state. Bar graphs
are easy, we just need one variable in our aes call, which will be our x axis. The
height of the bars will be the number of observations. We'll start with a basic plot
and make some adjustments as we go.

```{r}
ggplot(health_census, aes(StateDesc))+
  geom_bar()
```

So, that technically worked, but there are so many states that the x-axis is unreadable.
We can't tell what sate in here. We can fix this by adjusting the text in the x-axis. 

```{r}
ggplot(health_census, aes(StateDesc))+
  geom_bar()+
  # so this line tells ggplot to rotate the text in the x-axis by 90 degrees
  theme(axis.text.x = element_text(angle = 90))
```

This plot is a log better, now we can see the state names. But it's not awesome.
The states are ordered alphabetically, but that makes it difficult to compare states
to each other - do we have more rows from Maine or from New Mexico?

```{r}
# I add the fct_infreq function here to order by the counts. 
ggplot(health_census, aes(fct_infreq(StateDesc)))+
  geom_bar()+
  # so this line tells ggplot to rotate the text in the x-axis by 90 degrees
  theme(axis.text.x = element_text(angle = 90))
```

now we can see that Maine has slightly more observations than New Mexico.


# Part 4: Make a new plot using variables and chart type of your choice.


