---
title: "SOC 280 Lab Workshop 3"
output: html_notebook
---

Check out this cool data project and answer the questions:
https://googletrends.github.io/asne/?filter=gender&view=0

1. Interpret the plot as a whole, including the gender and race views.

2. Find the Chicago newsrooms. What do you think? Why is it like that?


Part 0: Review

0.1 Prepare your environment and load the airbnb data: make a new code chunk, load the library tidyverse, and then load the airbnb data into r. 

Part 1: Research Questions and Data Sources

Our Airbnb data comes from this website: http://insideairbnb.com/index.html

1.1 What is the purpose of this website? (that is, why did this person collect this data?)

1.2 Find an example on the website of a question or issue the authors used their data to answer or address. What were their findings?


Part 2:

Here's a sample research question:

What area of Chicago has the most expensive airbnb listings?

Before we answer the question, let's examine it.

2.1 What are the ideas that we'd need to measure to answer the question?

2.2 What columns from the airbnb dataset can we use to measure those ideas?

Bonus: Answer the research question.

Part 3:

New Research Question:

Does the location of an airbnb affect its price??

First, let's do an example.

I'm going to go through a few steps here to answer a version of this question. 

First, I want to look at a little less data. Instead of the whole airbnb set, I'm going to make a smaller data set with only a few varibles. To do so, I'm going to use a nifty trick: the pipe, or %>% (or |>). These three awesome characters allow me to take the output of one function and use it as the input for another function. Check it out:

```{r}
#new var              pipe! 
small_airbnb <- airbnb %>% select(price, neighbourhood, availability_365)
#this makes a new variable, 'small_airbnb' by 'selecting' the columns price, neighborhood, and availability

```

Check out the new variable to see that it only has two columns.

Ok, now that we've selected some columns, we can modify the data using the `mutate` function.

Mutate lets us make a new variable in our dataframe, often by using existing variables.
I want to make a new variable that just indicates whether there is any availaiblity 
in the next year. 

```{r}
small_airbnb <- small_airbnb %>% mutate(any_avail = availability_365 > 0)
```


Now, I'm going to limit it to units with at least some availability and limit to just a few neighborhoods. Just like before, I'll use the pipe, but now I'm using 'filter' instead of 'select.' 'select' let's me pick columns, 'filter' lets me filter out only the *rows* that match a certain rule. In this case, I'm matching two rules. First, I use '==' to match listings where my new column `any_avail` is TRUE. Then, I use the `%in%` operator to limit to listings with neighborhoods in the vector of neighborhood names I made.

```{r}
#first I make a list of neighborhood names that I want to keep          
hoods <- c("Near North Side", "West Town", "West Side", "Logan Square", "Loop", "South Chicago")
#then I use that list to filter only the rows that match my list: 
small_airbnb <- small_airbnb %>% filter(any_avail == TRUE, neighbourhood %in% hoods)
```


Now we can use one more new function combination, groupby + summarise, to get the mean price by neighborhood. This tells R to take our data, the put it into groups by neighborhood. Then, for each group, find the mean, or the average. Note that I add 'na.rm = TRUE' to the mean function. That is an extra argument that tells R to drop NA values. 

```{r}
#here I don't use a <- because I'm not saving the output as a variable, I'm just looking at it
small_airbnb %>% group_by(neighbourhood) %>% summarise(mean = mean(price, na.rm = TRUE))

``` 
##For each of these questions you'll need to write your own code. HOWEVER, the code you need is very similar to the code I used above. 

3.1 Use select to make a new dataset called 'medium_airbnb' with the following columns:
price, neighborhood, number_of_reviews

3.2 Filter your dataset to include only listings with more than 5 reviews


3.3 Use groupby and summarise to find the mean price (or total_price) by neighborhood


Bonus: Which neighborhood has the highest average total price?

