---
title: "SOC 225 Lab Workshop 6"
output: html_notebook
---

Check-In: 


Part 0: Prep 

0.1 We're going to use the dataset from Homework 3 today. So download homework 3 and load the airbnb_census file from it. It should have 9677 observations and 106 variables.  

0.2 This dataset includes a variable which indicates for each listing, which race has the most people in that zip code. For instance, my zipcode, 10025, has more white people than any other race. What is the name of this variable?

0.3 Donald Trump was born in Jamaica Hospital in zipcode 11418. What is the value of this variable for that zipcode?


Part 1: Titles and Labels

I realized we shoud have covered titles and labels before, so we're doing it now quickly.

Here's the distribution of prices in our dataset.
```{r}
ggplot(airbnb_census %>% select(clean_price) %>% drop_na())+ # I'm filtering out NAs to avoid warnings
  geom_histogram(aes(clean_price), bins=30) # I'm setting bins=30 for the same reason, you can mess with the bins too
```

Someone who wasn't familiar with this dataset might have trouble reading this plot though. We can add a title by just putting another plus and a title:
```{r}
ggplot(airbnb_census %>% select(clean_price) %>% drop_na())+
  geom_histogram(aes(clean_price), bins=30)+
  ggtitle("The Best Histogram of AirBnB Prices Ever")
```

That's better

We can also adjust the labels:

```{r}
ggplot(airbnb_census %>% select(clean_price) %>% drop_na())+
  geom_histogram(aes(clean_price), bins=30)+
  ggtitle("The Best Histogram of AirBnB Prices Ever")+
  xlab("Listing Price")+
  ylab("Number of Listings")
```

That looks solid now. Finally, I can improve data-ink ratio:
```{r}
mean =mean(airbnb_census$clean_price)
ggplot(airbnb_census %>% select(clean_price) %>% drop_na())+
  geom_histogram(aes(clean_price, alpha=.3), bins=30)+
  ggtitle("The Best Histogram of AirBnB Prices Ever")+
  xlab("Listing Price")+
  ylab("Number of Listings")+
  geom_vline(xintercept = mean(airbnb_census$clean_price, na.rm=TRUE), color='red')+
  theme_classic()+
  theme(legend.position="none")
```


1.1 What does the red line in the last plot mean? Check the code? Hint: bad pun.

1.2 The title isn't really that good, make it better.

1.3 Make a chart (type is up to you) using the number of bedrooms and properly label the axes.

Part 2: ggmap

There are lots of options for plotting geographic data in R. We're going to do (what I think is) the most simple way: using a map as a ggplot background. You'll need to install a package called 'ggmap'. Use the console to install it now, then load it in the codeblock below. (You'll see a logging output about google maps)

```{r}
library(ggmap)
```

Using ggmap, we usually first identify boundry latitude and longitude, and then plot the map:

```{r}
europe <- c(left = -12, bottom = 35, right = 30, top = 63) # rought lat/lon boundries for Europe
get_stamenmap(europe, zoom = 4, maptype = "toner-lite") %>% ggmap()  #pipe the get_stamenmap() function to ggmap()
```
  
You can also use different kinds of maptypes, and save the stamenmap to file if necessary:
  
```{r}
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
us_stamen <- get_stamenmap(us, zoom = 4, maptype = "watercolor") 
ggmap(us_stamen)
```
  
Read this code and see what it does. How does it set longitude and latitude? 
  
```{r}
new_york <- c(left = min(airbnb_census$longitude), 
              bottom =min(airbnb_census$latitude), 
              right = max(airbnb_census$longitude),
              top = max(airbnb_census$latitude))
nyc_stamen <- get_stamenmap(new_york, zoom = 10, maptype = 'toner-background')
ggmap(nyc_stamen)+
  geom_point(data=airbnb_census,aes(longitude, latitude, color=neighbourhood_group_cleansed))
```
  
Ggmap also works with google maps:
What's at the dot?
```{r}
get_googlemap("universityofwashington seattle", zoom = 12) %>% ggmap()+
  geom_point(aes(-122.2669889,47.6889121), size=5, color='purple')
```


You can see more here: https://github.com/dkahle/ggmap

2.1 Plot a stamen map of a place you'd like to go. 

2.2 Plot only listings with price greater than $600 on to a google map of New York.

2.3 Do you think New York is racially segregated? Could you make a plot that addressed that issue?

Extra Time? Extra Challenge

ET.1 Make a map plot that shows both price and neighborhood racial composition. 

ET.2 Download the airbnb data for Seattle (or another city of your choice) and make a map plot showing some information about the units.



