---
title: "SOC 280 LAB Homework 2"
author: "YOUR NAME HERE"
output: html_notebook
---
Write all code in the chunks provided!

Question one should be roughly analogous to what we've done in class. There are hints at the bottom of this document if you get stuck. If you still can't figure it out, go to google/stack exchange/ask a friend. Finally, email me or come to office hours :).



## 1.1: Piping Hot Variables

1.1.1 Set up your environment by: 
-A. Loading the tidyverse  
-B. Reading the 'health_census.csv' data
-C. Glimpse the health_census Data
  
```{r}

```
  
1.1.2 Use the data to answer this question: In how many of these tracts do more than 15% of people lack health insurance?
```{r}

```
  
1.1.3 Building on that work, what is the average percentage of people who are lacking health insurance in each state in the data? 
  
```{r}

```
  
## 1.2 Basic Visualization

We're going to answer a more complicated question now. What is the relationship between population density and high cholesterol.

The variable 'popden13_17' represents the tract population density; 'high_cholesterol' represents the percentage of people 18 and over who had high cholesterol.

1.2.1 Calculate and interpret the correlation between those variables.

```{r}

```

1.2.2 Make a visualization showing the relationship between 'popden13_17' and 'high_cholesterol'. Describe the plot.

```{r}

```

1.2.3 You might notice that the plot is... hard to interpret. That's partially because of the scale of the population density variable: most of the values are very low, but a significant portion of them are very high. With data like that, it's useful to apply an axis transformation. Remake your plot from 2.2 with an axis transformation, then interpret the plot.

```{r}

```

1.2.4 Perhaps you've noticed that the plot has a strange shape to it--it doesn't follow a single line, but seems to bend.
That's because there's a third variable influencing this relationship. It turns out that the relationship between population density and high cholesterol seems different in places with high proportions of white folks than in other places.

1.2.4a Use the `mutate` function to create a variabled 'high_white' that is TRUE when a neighborhood has a white population above the median for the data and FALSE when the population is below the median for the data.

```{r}

```

1.2.4b Remake the plot from 2.3 and set the color as the 'high_white' variable you created. Re-interpret this plot. Compare the interpretation of this plot to the plot you made in 2.2. Does this plot do a better job addressing the research question? Why or why not?

```{r}

```

# Part 2: Final Project Scaffolding


## Problem 2.1: Literature Review
*This question asks you to think deeply about the research question you're investigating. Each answer should be around 100 words.*

2.1.1: What dataset did you select (include a link again)? Why did you select it? What is your research question? What variables do you plan to use to answer your question?

2.1.2: Find at least two articles (at least one must be from an academic journal) that have addressed a question similar to your own. What data did they use? What problems did they have? *If you 'can't find' two articles, provide a screenshot of your search in the university library system from here: https://library.uic.edu*

2.1.3: What is one way that you have to modify or examine your data to begin to answer your question?

## Problem 2.2: Pipe your own data

2.2: Using the functions we've worked with in class (select, filter, mutate, groupby, summarise), plus any others you'd like to use, examine the key relationship from your research question.

You must:
a. Created a new dataset that only includes the variables you're interested in
b. Output a version of that dataset that filters it to certain values, hopefully ones you're interested in.
c. Create a mutated version of one of your variables (many of you will *need* to do this, but even if you don't, I want to see that you can)
d. Use groupby to group your data by one variable and see the mean (or similar) of another variable in those groups.

*Use as many codeblocks as you need*
  
```{r}

```


##Hints
1.2
Try using `count` function.

1.3
Ignore/Don't worry about NAs
You'll need to use group_by and summarize here.

2.2
a. use select()
b. use filter()
c. use mutate()
d. use groupby(var1) %>% summarise(mean = mean(var2))
