---
title: "SOC 225 Lab Workshop 5"
output: html_notebook
---

Check-In: 

Complete part 0. What's the population of my home zipcode, 10025?

Part 0: Review (Same as Workshop 4, cut and paste~)

0.1 Load the tidyverse and the 'airbnb_census.csv' dataset. This dataset includes all of the listings from the full airbnb dataset that include square feet. It also has some information from the 2010 census: total population and number of white, black, asian, and latinx folks in the neighborhood.

0.2 We're going to need a clean price variable for this lab, so take the airbnb dataset and create a clean price variable, it should be numeric.

0.3 We'll also need some racial proportion variables. The following code makes a 'white_proportion' variable by dividing the white population by the total population. Do the same for the other race variables.

```{r}
airbnb <- airbnb %>% mutate(
  white_proportion = white/full_pop,
  clean_price = as.numeric(gsub('\\$','',price))
)
```


Part 1: Variation

We can use histograms to look at the variation of continuous variables.

Here's the distribution of our square_feet variable
```{r}
ggplot(airbnb)+
  geom_histogram(aes(square_feet))
```

I can compare the distribution of square footage by area:

```{r}
ggplot(airbnb, aes(x = square_feet, colour = neighbourhood_group_cleansed)) +
  geom_freqpoly()
```

But this doesn't do a good job of comparing the neighborhoods because it uses total counts. The following sets the y axis to density and uses a facet wrap.

```{r}
ggplot(airbnb, aes(x = square_feet, y=..density..)) +
  geom_freqpoly()+
  facet_wrap(~neighbourhood_group_cleansed)
```

1.1 Repeat the above analyis but use each of the race variables instead of square feet. What do these visualizations tell you?

1.2 Use the information from Workshop 4 to make a bar graph of guests included. What does this visualization tell you?

1.Bonus Use the mapping technique we used last time to re-create the lat-lon plot of the units, but trying coloring them by racial proportions. What are you creating when you do this? What does this tell you about New York?

Part 2: Covariation

What visualizations are really good at, hwoever, is comparing how two variables vary together. We already saw that a bit when we used scatterplots.

Let's see how price varies with square footage.

```{r}
ggplot(airbnb %>% filter(square_feet>10))+
  geom_point(aes(square_feet,clean_price))
```

This looks similar to what we saw before. 

2.1 Repeat the above comparision but compare the relationship between white proportion and black proportion with price. What do your visualizations tell you?

2.2 Make a new variable that is 1 if a zipcode is majority white and 0 if it is less than 50% white. Make a boxplot comparing the prices in the two neighborhoods.

2.Bonus Make a visualization that shows how the relationship between square feet and price changes depending on the racial composition of the neighborhood.

Extra Time? Extra Challenge

ET.1 Estimate the racial composition for each neighborhood (not neighborhood group)

ET.2 Using neighborhoods as your unit of analyisis, plot the relationship between neighborhood racial composition and neighborhood average price.


