---
title: "SOC 225 Lab Workshop 4"
output: html_notebook
---

Check-In: 

Part 0: Review

0.1 Note that I haven't included a /data file or a copy of the airbnb data. Import the tidyverse, download the data, put it in the data file, and load it... All that good stuff.

0.2 We're going to need a clean price variable for this lab, so take the airbnb dataset and create a clean price variable, it should be numeric. You might need to divide by 100... Call it 'clean_price' or the my example code won't work :). 

0.Bonus Make a new variable that is the number of characters in the 'name' 

Part 1:

We'll start with an example:

Some questions are best answered visually. Let's use this one:

How is listing price related to square footage?
*take a moment to make a hypothesis*

First, without a visualizaiton, I could look at a correlation:
```{r}
#first make a small df to include only the variables I want, then drop NA values and filter to only listings where square_feet is greater than 10 (because they're probably fake)
airbnb_small <- airbnb %>% select(clean_price, square_feet, bedrooms, bathrooms, neighbourhood_group_cleansed) %>% drop_na() %>% filter(square_feet>10)
# do a simple correlation between my variables
cor(airbnb_small$clean_price,airbnb_small$square_feet)
```
This result shows that as price increases, squarefootage tends to increase and vice versa. The correlation is moderately strong. 

How could we answer this question using visualizations? 

In this situation, though, we still don't have a good understanding of this relationship. It might help to make a graph.

```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price))
```

1.1 What is the correlation between number of bedrooms and number of bathrooms?

1.2 Plot the relationship.

1.Bonus: Graph the relationship between price and number of characters.

Part 2: Research Questions and Data Sources

What makes a good visualization?

Check out these resources to answer the following questions.

The Double Think: https://thedoublethink.com/tuftes-principles-for-visualizing-quantitative-information/

Darkhorse: https://www.darkhorseanalytics.com/blog/data-looks-better-naked

2.1 What is the data-ink ratio?

2.2 What is chart junk?

2.3 Look at the example chart on the darkhorse website. Use some specific examples to describe how they improve their visualization.

2.4 Which change from 1.3 do you think is most effective?

2.Bonus Modify my code from Part 1 to make a better visualization.



Part 3: Different Visualization Types

Let's consider our square footage graph from before:

```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price))
```

There are two points with very high prices, over $100,000 per night. Where do you think these units are?

Uisng 'color' in our aes(), we can add colors and labels to our visualization:

```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price, color=neighbourhood_group_cleansed))
```

3.1 Where are the most expensive units? Where are the largest units?

We can also change the size of the dots

```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price, size=bedrooms))
```

Now dots are sized by the number of bedrooms. 

Sometimes it's nice to have more than one graph. Our graph which used the colors for the bouroughs in New York wasn't very effective because the dots overlapped. We can use facet wrap to make a bunch of plots of similar data.

```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price))+
  facet_wrap(~ neighbourhood_group_cleansed, nrow = 2)
```

Now we can see differences in area a bit more clearly. Poor Staten Island~

We can also add trendlines:
```{r}
ggplot(data = airbnb_small) + 
  geom_point(mapping = aes(x = square_feet, y = clean_price))+
  geom_smooth(mapping = aes(x = square_feet, y = clean_price), method = 'glm', se=FALSE)
```

There are some other kinds of plots we can make:

Histograms:

```{r}
# I can use dplyr right in the ggplot! 
# This is a histogram including only units with prices under $1,000 a night 
# You can try taking out the filter and see what happens.
ggplot(airbnb %>% filter(clean_price<1000))+
  geom_histogram(aes(clean_price), bins=35)
```


Box and whisker plots:
```{r}
# I can even start with a pipe and then pipe it into ggplot!
# Here I take airbnb, filter it to less than 7 bedrooms and less than $2,500 a night
airbnb %>% filter(bedrooms<7, clean_price<2500) %>%
  ggplot()+
  # and then plot a boxplot
    geom_boxplot(aes(factor(bedrooms), clean_price))
```

Bar graphs:

```{r}
ggplot(airbnb)+
  geom_bar(aes(neighbourhood_group_cleansed))
```

3.2 Change the above code so that we can see the distribution of units which cost more than $5,000 a night. Less than $100 a night?

Plots with longitude and latitude:

```{r}
ggplot(airbnb) +
  geom_point(aes(longitude, latitude, color=neighbourhood_group_cleansed)) +
  coord_quickmap()
colnames(airbnb)
```

3.3 Modify the above to show color by number of bedrooms, by price.

3.4 Make a new plot using variables and chart type of your choice.

Extra time?

Answer these questions with plots:
E.1 Do more expensive units get reviewd more often?
E.2 Are units that have been on AirBnb longer more expensive?
E.3 Do those results differ by neighborhood group?

#Hints!
0.Bonus try the nchar() function
