---
title: "SOC 225 Lab Workshop 3"
output: html_notebook
---

Check-In: 

Check out this cool data project and answer the questions on the board:
https://googletrends.github.io/asne/?filter=gender&view=0


Part 0: Review

0.1 Note that I haven't included a /data file or a copy of the airbnb data. See if you can download that data, put it into the right directory, make a new code chunk, load the library tidyverse, and then load the airbnb data into r. 

Part 1: Research Questions and Data Sources

Our Airbnb data comes from this website: http://insideairbnb.com/index.html

1.1 What is the purpose of this website? (that is, why did this person collect this data?)

1.2 Find an example on the website of a question or issue the authors used their data to answer or address. What were their findings?

Bonus: Can you verify any of their results with the data that you have (remember that we're only using a subset of the full data in the lab, so the exact numbers might not match)


Part 2:

Here's a sample research question:

What area of New York has the most expensive airbnb listings?

Before we answer the question, let's examine it.

2.1 What are the ideas that we'd need to measure to answer the question?

2.2 What columns from the airbnb dataset can we use to measure those ideas?

Bonus: Answer the research question.

Part 3:

New Research Question:

Does the location of an airbnb affect its price??

First, let's do an example.

I'm going to go through a few steps here to answer a version of this question. 

First, I want to look at a little less data. Instead of the whole airbnb set, I'm going to make a smaller data set with only a few varibles. To do so, I'm going to use a nifty trick: the pipe, or %>%. These three awesome characters allow me to take the output of one function and use it as the input for another function. Check it out:

```{r}
#new var              pipe! 
small_airbnb <- airbnb %>% select(price, neighbourhood_cleansed)
#this makes a new variable, 'small_airbnb' by 'selecting' the columns price and neighborhood_cleansed

```

Check out the new variable to see that it only has two columns.

Now, I'm going to limit the set to just a few neighborhoods. Just like before, I'll use the pipe, but now I'm using 'filter' instead of 'select.' 'Select' let's me pick columns, 'filter' lets me filter out only the *rows* that match a certian rule.

```{r}
#first I make a list of neighborhood names that I want to keep          (my home 'hood!)
hoods <- c("Harlem", "Bushwick", "SoHo", "Willamsburg", "East Village", "Upper West Side")
#then I use that list to filter only the rows that match my list: 
small_airbnb <- small_airbnb %>% filter(neighbourhood_cleansed %in% hoods)
```

I have another problem: that pesky price variable which isn't a number, but a string. Let's take care of that using 'mutate.'

```{r}
small_airbnb <- small_airbnb %>% mutate(clean_price = as.numeric(gsub('\\$|,','',price)))
```

Finally, we can use one more new function combination, groupby + summarise, to get the mean price by neighborhood.

```{r}
#here I don't use a <- because I'm not saving the output as a variable, I'm just looking at it
small_airbnb %>% group_by(neighbourhood_cleansed) %>% summarise(mean = mean(clean_price))

``` 
##For each of these questions you'll need to write your own code. HOWEVER, the code you need is very similar to the code I used above. 

3.1 Use select to make a new dataset called 'medium_airbnb' with the following columns:
zipcode, price, cleaning_fee, and number_of_reviews

3.2 Filter your dataset to inlcude only listings with more than 5 reviews

3.3 Clean the price variable into a 'clean_price' variable. (Bonus: clean the cleaning fee variable too and make a new variable called 'total_price' which is the sum of clean_price and clean_cleaning_fee)

3.3 Use groupby and summarise to find the mean price (or total_price) by zipcode


Bonus: Which zipcode has the highest average total price?


Extra time? Try the tidycensus package.

E.1 Install the tidycensus package
Info here:https://github.com/walkerke/tidycensus

E.2 Use the tidy census to get information about race by zipcode in New York State.
Use the get_decennial() function
Set geography to 'zcta'
Census variables for race: 
White = "P0050003"
Black = "P0050004"
Asian = "P0050006"
Hispanic = "P0040003"
summary_var = "P0010001" (full population)

E.3 Reshape the resulting data into a format that you can merge with the Airbnb data and see what you can see.
